Supplement 1: Testing different distributions

### 2018-09-14 pkg/disttree/inst/draft/supplement/supplement_distributions.Rnw
revised version of supplement 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2282 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 01c0f077e4047722b3019c3d9da67a06e3535524

\documentclass[nojss]{jss}

@@

@@

The distribution can not only be exchanged for the distributional forest 
but for all the models evaluated in this application:

@@

\item \emph{EMOS:} Standard ensemble model output statistics
  models use the ensemble mean of total precipitation as regressor in the location
  submodel and the corresponding ensemble standard deviation in the scale submodel.
  The parameters are estimated by maximum likelihood, using an identity
  link for the location part and a log link for the scale part (following the advice of
  \cite{Gebetsberger+Messner+Mayr:2017}).

@@

\item \emph{Boosted GAMLSS:} Smooth additive splines are selected
  automatically from all available variables, using non-cyclic boosting
  for parameter estimation (\cite{Hofner+Mayr+Schmid:2016, Messner+Mayr+Zeileis:2017}).
  This updates the predictor terms for the location or scale submodels iteratively
  by maximizing the log-likelihood only for the variable yielding the
  biggest improvement. The iteration stops early -- before fully maximizing
  the in-sample likelihood -- based on a (computationally intensive)
  out-of-bag bootstrap estimate of the log-likelihood.

@@

\begin{align}
f_{\textit{logistic}}(Y; \mu, \sigma) &= 
     \frac{\exp{\left(-\frac{Y - \mu}{\sigma}\right)}}
     {\sigma \cdot \left(1 + \exp{\left(-\frac{Y - \mu}{\sigma}\right)}\right)^2} \\
F_{\textit{logistic}}(Y; \mu, \sigma) &=
     \frac{1}{1+ \exp{\left(\frac{\mu}{\sigma}\right)}}.
\end{align}

Regardless of the chosen distribution it has to be considered 
that precipitation data usually includes a point mass at zero.
In order to deal with this point mass one of two common strategies 
can be used.
\begin{itemize}
\item Censored distribution (\cite{Winkelmann:2006}): 
In general, censoring data usually leads 
to a point mass at the censoring point which has to be considered in 
the modeling process. Therefore, censored distributions are applied as 
they are able to capture this point mass. 
In precipitation data, observations can not be negative, thus, left-censoring
it at zero does not change the data. However, there is a naturally given
point mass at zero (days without rain) making a censored distribution an 
appropriate choice to model precipitation data.
For a chosen distribution with probability density function $f$ and
distribution function $F$ the log-likelihood function of the corresponding 
censored distribution is
\begin{equation}
\ell_{\textit{cens}}(\mu, \sigma; Y) = 
\begin{cases}
    \log\left\{f(Y; \mu, \sigma) \right\}, & \text{if } Y > 0 \\[0.2cm]
    \log\left\{F(0; \mu, \sigma) \right\}, & \text{if } Y = 0
\end{cases}
\end{equation}
as seen in Equation~2.1.1. of the paper for a Gaussian distribution.

\item Hurdle model / Two-part model (\cite{Cragg:1971}, \cite{Mullahy:1986}): 
An additional parameter $\nu$ describing the probability that there is 
any precipitation at all, 
i.e. $\nu = \mathbb{P}(Y>0)$, is estimated by a binary model. 
Then, a second model is fit to the subset of the data containing 
only positive observations applying a left-truncated at zero 
distribution to estimate the location parameter $\mu$ and scale 
parameter $\sigma$ of the underlying latent variable.
The three-parametric log-likelihood function for the full hurdle model 
with parameter vector $\bm{\theta} = (\mu, \sigma, \nu)$ is
\begin{equation} \label{eq:hurdle}
\ell_{\textit{hurdle}}(\mu, \sigma, \nu; Y) = 
\begin{cases}
    \log\left\{\nu \cdot \frac{f(Y; \mu, \sigma)}{1 - F(0; \mu, \sigma)} \right\}, & \text{if } Y > 0 \\[0.2cm]
    \log(1 - \nu), & \text{if } Y = 0.
\end{cases}
\end{equation}

@@

\section{Stations}
Out of the 95 observation stations 18 are considered for this 
benchmark study (see Figure~\ref{fig:map}). 
They have been selected based on the region
they are in, their altitude and the performance of the 
different models in the application in the paper such that
all of these features are represented equally.

\begin{figure}[h]
\vspace*{-3cm}
<<map, fig=TRUE, echo=FALSE>>=
##  map
data("StationsTyrol", package = "RainTyrol")
data("MapTyrol", package = "RainTyrol")
library("sp")
sp <- SpatialPointsDataFrame(subset(StationsTyrol, select = c(lon, lat)),
                             data = subset(StationsTyrol, select = -c(lon, lat)),
                             proj4string = raster::crs(MapTyrol$RasterLayer))

plot(MapTyrol$SpatialPolygons)
points(sp, pch = 21, col = "black", las = 1, cex = 1)
points(sp[c(5, 6, 16, 20, 23, 25, 32, 33, 46, 47, 48, 56, 57, 70, 71, 79, 82, 83),], pch = 21, bg = hcl(325, 100, 70), cex = 1.5)
@
\vspace*{-3cm}
\caption{\label{fig:map}Map of Tyrol with all 95 observation stations and the 18 considered stations highlighted in pink.}
\end{figure}

\section{Results}

@@

xyplot(crps ~ distribution | station, groups = ~ method, 
       data = means, auto.key = TRUE, 
       type = "o", lwd = 2, lty = 1, layout = c(3,6))
@
\caption{\label{fig:lattice}CRPS values of all 4 methods, once applied with each of the three distributions for each of the 18 observation stations.}
\end{figure}

@@

@
\caption{\label{fig:box_crps}(Mean) CRPS values of each of the 18 observation stations, separated by method and distribution.}
\end{figure}

  par(mfrow = c(1,3))
  boxplot(crps_ss_method~method, data = means_sel, 
          subset = (distribution == "gaussian"), main = "gaussian",
          ylim = c(-0.12,0.23), axes = FALSE, ylab = "CRPS skill score")
  axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 2)
  axis(1, 0:4, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS", ""), las=2)
  axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_method~method, data = means_sel, 
          subset = (distribution == "logistic"), main = "logistic",
          ylim = c(-0.12,0.23), axes = FALSE)
  axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 2)
  axis(1, 0:4, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS", ""), las=2)
  axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_method~method, data = means_sel, 
          subset = (distribution == "hgaussian"), main = "hgaussian",
          ylim = c(-0.12,0.23), axes = FALSE)
  axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 2)
  axis(1, 0:4, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS", ""), las=2)
  axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
  abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_EMOS}(Mean) CRPS skill score values of each of the 18 observation stations, separated by method and distribution. EMOS is chosen as reference model.}
\end{figure}

  par(mfrow = c(1,4))
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "Distributional forest"), main = "Distributional forest",
          ylim = c(-0.12,0.07), las = 2, ylab = "CRPS skill score")
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "Prespecified GAMLSS"), main = "Prespecified GAMLSS",
          ylim = c(-0.12,0.07), las = 2)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "Boosted GAMLSS"), main = "Boosted GAMLSS",
          ylim = c(-0.12,0.07), las = 2)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "EMOS"), main = "EMOS",
          ylim = c(-0.12,0.07), las = 2)
  abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_gaussian}(Mean) CRPS skill score values of each of the 18 observation stations, separated by method and distribution. Gaussian is chosen as reference distribution.}
\end{figure}

\section{Discussion}
As explained above the the parameter $\nu$ is modeled separately from $\mu$ and $\sigma$ 
in the GAMLSS framework which is why two separate models can be built. For the forest model
an alternative approach would be to build just one model employing the three-parametric
distribution family. However, testing both versions of the forest model in this application study shows that they both perform almost equally. Over the 18 considered stations the 
absolute difference in terms of CRPS between the two-forest model and the three-parametric forest is on average 0.0078 and lies between 0.0014 and 0.0267 for each station. Comparing
these values to the mean CRPS values in Figure~\ref{fig:box_crps} lying between 0.6 and 1.2
confirms the relatively small difference between the two versions of the forest model.

<<compare_forests, results=hide, echo=FALSE>>=
  load("crps_forests.rda")
  sum(crps_forests[,"forest_2p"] > crps_forests[,"forest_h"])
  summary(crps_forests[,"forest_h"] - crps_forests[,"forest_2p"])
  summary(abs(crps_forests[,"forest_h"] - crps_forests[,"forest_2p"]))
  colMeans(crps_forests)  
@

\bibliography{ref_supplement_distributions.bib}

>>>>

%\documentclass[a4paper,11pt]{article}
\documentclass[nojss, shortnames]{jss}

@@

\Abstract{In the application discussed in the paper ''Distributional 
Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain''
the underlying distribution is assumed to a be left-censored at zero Gaussian distribution.
In this suppplement the exact same evaluations as in the paper are performed employing
other distribution assumption. In particular, a left-censored at zero logistic distribution 
and a hurdle-model are applied and their results are compared to those presented in the paper.}

\Keywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, GAMLSS}

\Address{
  Lisa Schlosser, Reto Stauffer, Achim Zeileis \\
  Universit\"at Innsbruck \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Lisa.Schlosser@uibk.ac.at}, \email{Reto.Stauffer@uibk.ac.at},\\
  \email{Achim.Zeileis@R-project.org} \\
  URL: \url{https://www.uibk.ac.at/statistics/personal/schlosser-lisa/},\\
  \phantom{URL: }\url{https://retostauffer.org/},\\
  \phantom{URL: }\url{https://eeecon.uibk.ac.at/~zeileis/}\\

  Torsten Hothorn\\
  Universit\"at Z\"urich \\
  Institut f\"ur Epidemiologie, Biostatistik und Pr\"avention \\
  Hirschengraben 84\\
  CH-8001 Z\"urich, Switzerland \\
  E-mail: \email{Torsten.Hothorn@R-project.org}\\
  URL: \url{http://user.math.uzh.ch/hothorn/}\\

}

@@

To compare the performance of the distributional forest to other commonly used methods the exact same models as in the paper are evaluated on the same data set and applying the same settings while only the distribution is exchanged. Further details on the models can be found in Section~3.2.. In particular, Table~2 gives an overview over the models and their specifications. As a brief reminder the approaches are summarized below.

@@

\item \emph{EMOS} \citep{Gneiting+Raftery+Westveld:2005}: 
Standard ensemble model output statistics models use the ensemble 
mean of total precipitation as regressor in the location submodel 
and the corresponding ensemble standard deviation in the scale submodel.
The parameters are estimated by maximum likelihood.

@@

\item \emph{Boosted GAMLSS:} Smooth additive splines are selected
automatically from all available variables, using non-cyclic boosting
for parameter estimation (\cite{Hofner+Mayr+Schmid:2016, Messner+Mayr+Zeileis:2017}).
The optimal stopping point for the iteration is evaluated based on 
a (computationally intensive) out-of-bag bootstrap estimate of the log-likelihood.

@@

\begin{align}
f_{\textit{logistic}}(Y; \mu, \sigma) &= 
     \frac{\exp{\left(-\frac{Y - \mu}{\sigma}\right)}}
     {\sigma \cdot \left(1 + \exp{\left(-\frac{Y - \mu}{\sigma}\right)}\right)^2} \\
F_{\textit{logistic}}(Y; \mu, \sigma) &=
     \frac{1}{1+ \exp{\left(-\frac{Y - \mu}{\sigma}\right)}}.
\end{align}

Both distributions have to be adapted in our application to capture that
precipitation is non-negative with a point mass at zero (= no rain).
Two common strategies for this are to employ either \emph{left-censoring at zero}
or a \emph{hurdle} (or two-part) approach, combining a binary (zero vs.\ greater) and
a truncated (postive) part \citep[see][for an overview]{Winkelmann:2006}.
\begin{itemize}
\item Censored distribution \citep{Winkelmann:2006}: 
%In general, censoring data usually leads 
%to a point mass at the censoring point which has to be considered in 
%the modeling process. Therefore, censored distributions are applied as 
%they are able to capture this point mass. 
%In precipitation data, observations can not be negative, thus, left-censoring
%it at zero does not change the data. However, there is a naturally given
%point mass at zero (days without rain) making a censored distribution an 
%appropriate choice to model precipitation data.
For a chosen distribution with probability density function $f$ and
distribution function $F$ the log-likelihood function of the corresponding 
censored distribution is
\begin{equation}
\ell_{\textit{cens}}(\mu, \sigma; Y) = 
\begin{cases}
    \log\left\{f(Y; \mu, \sigma) \right\}, & \text{if } Y > 0 \\[0.2cm]
    \log\left\{F(0; \mu, \sigma) \right\}, & \text{if } Y = 0
\end{cases}
\end{equation}
as seen in Equation~2.1.1. of the paper for a Gaussian distribution.
For positive values of $Y$ the censored version of the density function
equals the original non-censored version. For $Y=0$ the censored density
function takes the values of the probability mass lying left of the 
censoring point. Therefore, the parameters $\mu$ and $\sigma$ control
both parts, the point mass and the positive part of the distribution.

\item Hurdle model / Two-part model (\cite{Cragg:1971}, \cite{Mullahy:1986}): 
An additional parameter $\nu$ describing the probability that there is 
any precipitation at all, i.e. $\nu = \mathbb{P}(Y>0)$, is included in the model.
%Then, a second model is fit to the subset of the data containing 
%only positive observations applying a left-truncated at zero 
%distribution to estimate the location parameter $\mu$ and scale 
%parameter $\sigma$ of the underlying latent variable.
The three-parametric log-likelihood function for the full hurdle model 
with parameter vector $\bm{\theta} = (\mu, \sigma, \nu)$ is
\begin{equation} \label{eq:hurdle}
\ell_{\textit{hurdle}}(\mu, \sigma, \nu; Y) = 
\begin{cases}
    \log\left\{\nu \cdot \frac{f(Y; \mu, \sigma)}{1 - F(0; \mu, \sigma)} \right\}, & \text{if } Y > 0 \\[0.2cm]
    \log(1 - \nu), & \text{if } Y = 0.
\end{cases}
\end{equation}
For $Y=0$ the probability that it won't rain is expressed by $1-\nu$. For $Y>0$ the probability that it will rain $\nu$ is multiplied with the left-truncated at zero density function. Contrary to the censored model, here it is only the parameter $\nu$ controlling the point mass while $\mu$ and $\sigma$ only influence the distribution of the positive observations.
\end{itemize}

@@

\section{Stations}
Out of the 95 observation stations 15 are considered for this 
benchmark study (see Figure~\ref{fig:map}). 
They have been selected based on the region
they are in, their altitude and the performance of the 
different models in the application in the paper such that
all of these features are represented equally.

\begin{figure}[h]
\vspace*{-3cm}
\setkeys{Gin}{width=0.99\textwidth}
<<map_stationwise, fig=TRUE, echo=FALSE>>=
##  map
data("StationsTyrol", package = "RainTyrol")
data("MapTyrol", package = "RainTyrol")
library("sp")
sp <- SpatialPointsDataFrame(subset(StationsTyrol, select = c(lon, lat)),
                             data = subset(StationsTyrol, select = -c(lon, lat)),
                             proj4string = raster::crs(MapTyrol$RasterLayer))

plot(MapTyrol$SpatialPolygons)
points(sp, pch = 21, col = "darkgrey", las = 1, cex = 0.6)
points(sp[c(5, 6, 20, 23, 25, 32, 33, 46, 47, 56, 57, 70, 79, 82, 83),], pch = 21, bg = hcl(325, 100, 70), cex = 1.5)
# stationname beneath
text(sp[c(25, 47, 57, 70),], 
     labels = StationsTyrol$name[c(25, 47, 57, 70)],
     pos=1, cex=0.75)
# stationname left
text(sp[c(20, 32, 46, 79, 83),], 
     labels = StationsTyrol$name[c(20, 32, 46, 79, 83)],
     pos=2, cex=0.75)
# stationname above
text(sp[c(6, 33, 82),], 
     labels = StationsTyrol$name[c(6, 33, 82)],
     pos=3, cex=0.75)
# stationname right
text(sp[c(5, 23, 56),], 
     labels = StationsTyrol$name[c(5, 23, 56)],
     pos=4, cex=0.75)
@
\vspace*{-3.5cm}
\caption{\label{fig:map}Map of Tyrol with all 95 observation stations and the 15 considered stations highlighted in pink.}
\end{figure}

\section{Results}
In Figure~\ref{fig:lattice} each panel belongs to one of the 15 observation stations. The three distributions (left-censored at zero Gaussian distribution = \textit{gaussian}, hurdle model based on the Gaussian distribution = \textit{hgaussian}, left-censored at zero logistic distribution = \textit{logistic}) are compared based on the resulting CRPS values for each of the applied methods that are represented by different colours and symbols. Overall it can be stated that the forest model performs almost equally well with all three distributions. Therefore it seems to be very robust against missspecifications of the distribution.
For the two GAMLSS models the censored Gaussian and the hurdle model perform similarly at most of the station while applying the censored logistic distribution leads to an increase in CRPS, i.e. to a decline in performance, especially for the prespecified GAMLSS.
For the EMOS model exchanging the distribution does not have a strong effect on its performance, similar to the forest model. In particular, for the two censored distributions the results are almost equal, only for the hurdle model the CRPS values vary slightly, however in both directions, depending on the station.

When applying a hurdle model for the stations Oetz and Innvervillgraten the prespecified GAMLSS computationally failed to build/fit a model. Therefore this method is not represented in Figure~\ref{fig:lattice} for these two cases.

In stations like Ladis-Neuegg, Matrei in Osttirol or Oetz all methods seem to be very robust and stable over varying distributions. The strongest influence of an exchange of the distribution can be seen for example at stations Jungholz and St.Johann im Walde for the prespecified GAMLSS.

In Figure\ref{fig:box_crpss_gaussian} the tested distributions are compared for each method based on CRPS skill score with the censored Gaussian distribution as reference represented by the horizontal line at zero. The hurdle model leads to a slight improvement for the forest model and the prespecified GAMLSS. Following the median this improvement is a bit stronger for the boosted GAMLSS, however, with a higher deviation as represented by the larger size of the box, similarly as for the EMOS, however with a negative influence indicated by the negative median. Choosing a logistic distribution instead of a Gaussian clearly reduces the strength of the two GAMLSS models, while it has almost no influence on the forest and the EMOS models.

Alltogether it can be stated that the left-censored at zero Gaussian distribution seems to be an appropriate choice while the other two distributions are possible and mostly also reasonable alternatives. However, they do not lead to an overall improvement in performance for this application.

@@

xyplot(crps ~ distribution | station, groups = ~ method, 
       data = means, auto.key = TRUE, 
       type = "o", lwd = 2, lty = 1, layout = c(3,6))
@
\caption{\label{fig:lattice}CRPS values of all 4 methods, once applied with each of the three distributions for each of the 15 observation stations.}
\end{figure}

  par(mfrow = c(1,4))
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "Distributional forest"), main = "Distributional forest",
          ylim = c(-0.12,0.07), las = 2, ylab = "CRPS skill score")
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "Prespecified GAMLSS"), main = "Prespecified GAMLSS",
          ylim = c(-0.12,0.07), las = 2)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "Boosted GAMLSS"), main = "Boosted GAMLSS",
          ylim = c(-0.12,0.07), las = 2)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "EMOS"), main = "EMOS",
          ylim = c(-0.12,0.07), las = 2)
  abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_gaussian}(Mean) CRPS skill score values of each of the 15 observation stations, separated by method and distribution. Gaussian is chosen as reference distribution.}
\end{figure}

  par(mfrow = c(1,3))
  boxplot(crps_ss_method~method, data = means_sel, 
          subset = (distribution == "gaussian"), main = "gaussian",
          ylim = c(-0.12,0.23), axes = FALSE, ylab = "CRPS skill score")
  axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 2)
  axis(1, 0:4, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS", ""), las=2)
  axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_method~method, data = means_sel, 
          subset = (distribution == "logistic"), main = "logistic",
          ylim = c(-0.12,0.23), axes = FALSE)
  axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 2)
  axis(1, 0:4, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS", ""), las=2)
  axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_method~method, data = means_sel, 
          subset = (distribution == "hgaussian"), main = "hgaussian",
          ylim = c(-0.12,0.23), axes = FALSE)
  axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 2)
  axis(1, 0:4, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS", ""), las=2)
  axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
  abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_EMOS}(Mean) CRPS skill score values of each of the 15 observation stations, separated by method and distribution. EMOS is chosen as reference model.}
\end{figure}

@@

@
\caption{\label{fig:box_crps}(Mean) CRPS values of each of the 15 observation stations, separated by method and distribution.}
\end{figure}

\section{Discussion}
As mentioned by \cite{Hofner+Mayr+Schmid:2016} the methods provided by GAMLSS ``can be unstable, especially when it comes to selecting possibly different sets of variables for multiple distribution parameters''. These problems have also occured in a few situations for the 
prespecified GAMLSS model within the cross-validation. In case of this method failing to 
build a model it is represented by \code{NA}s in the evaluation.

As explained above the the parameter $\nu$ is modeled separately from $\mu$ and $\sigma$ 
in the GAMLSS framework which is why two separate models can be built. For the forest model
an alternative approach would be to build just one model employing the three-parametric
distribution family. However, testing both versions of the forest model in this application study shows that they both perform almost equally. Figure~\ref{fig:box_crpss_forests} shows
the CRPS skill score of the three-parametric forest model over the considered 15 observation
stations with the two-part forest model as reference. Even though this alternative version
of the forest model seems to perform better on average, according to the median this 
improvement is not even one percent, hence neglectably small. Therefore, the two foreset version can be stated to perform almost equally.

\begin{figure}[h]
\center
\setkeys{Gin}{width=0.5\textwidth}
<<compare_forests, fig=TRUE, results=hide, echo=FALSE, width=4, height=4>>=
  load("crps_forests.rda")
  crps_ss_forest_2p <- data.frame(1 - crps_forests[,"forest_2p"] / crps_forests[,"forest_h"])
  names(crps_ss_forest_2p) <- "Three-parametric forest"
  boxplot(crps_ss_forest_2p, ylab = "CRPS skill score", names = c("Three-parametric forest"))
  abline(h = 0, col = pal[5], lwd = 2)
  axis(1, 0:2, c("","Three-parametric forest", ""), las=1)
  #sum(crps_forests[,"forest_2p"] > crps_forests[,"forest_h"])
  #summary(crps_forests[,"forest_h"] - crps_forests[,"forest_2p"])
  #summary(abs(crps_forests[,"forest_h"] - crps_forests[,"forest_2p"]))
  #colMeans(crps_forests)  
@
\caption{\label{fig:box_crpss_forests}(Mean) CRPS skill score values of each of the 15 observation stations for the three-parametric forest model with the two part forest model as reference model.}
\end{figure}

%bibliographystyle{jss}
\bibliography{ref_supplement.bib}
